An automatic approach to find all possible calendar effects (periodic patterns) for an univariate time series

### 2021-02-24 paper/main.Rmd
added example wpds

```{r null4by2,fig.cap=" A graphical display with two categories mapped to x-axis and 4 categories mapped to facets with the distribution of a continuous random variable plotted on the y-axis. Display a is not interesting as the distribution of the continuous rv does not depend across x-axis or facet categories. Display b and c are more interesting than a since there is a change in distribution either across facets(b) or x-axis(a). Display d is most interesting as distribution of the rv changes across both facet and x-axis variable.", eval = FALSE}

# sim_varf_dist1 = function(nx, nfacet){
#   rep(dist_normal(seq(mean,mean*nfacet, by = mean), sd), each = nx)}

p1 <- sim_panel(nx = 2, nfacet = 3, ntimes = 500) %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot()

simulated_data <- sim_panel(nx = 2,
          nfacet = 3,
          ntimes = 500,
          sim_dist = rep(dist_normal(seq(5, 15, 5), 5), each = 2))

p2 <- simulated_data %>% ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot()

p3 <- simulated_data %>% ggplot(aes(x = as.factor(id_facet), y = sim_data)) + facet_wrap(~id_x) + geom_boxplot()

p4 <- sim_panel(nx = 2,
          nfacet = 3,
          ntimes = 500,
          sim_dist = dist_normal(seq(5,30, 5), 5)) %>% 
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot()

ggpubr::ggarrange(p1, p2, p3, p4, nrow = 2, ncol = 2,
                  common.legend = TRUE,
                  labels = c("a", "b", "c", "d"))

@@

*incorporate from the document combining_normalisation_method.Rmd*

@@

@@

```{r varall}

@@

# plot
p_varall <- sim_panel_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
  ggtitle(paste("Design $D_{var_{all}}$:", varall))
```

```{r varx}
sim_varx_normal <- function(nx, nfacet, mean, sd, w) {
   rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w), sd), nfacet)
 }

sim_panel_varx <- sim_panel(
   nx = 2, nfacet = 3,
   ntimes = 5,
   sim_dist = sim_varx_normal(2, 3, 0, 1, 10)
 ) %>% unnest(data)

@@

# plot
p_varx <- sim_panel_varx %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle(paste("Design $D_{var_x}$:", varx))
```

```{r varf}
sim_varf_normal <- function(nx, nfacet, mean, sd, w) {
  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w), sd), each = nx)
}
sim_panel_varf <- sim_panel(
  nx = 2, nfacet = 3,
  ntimes = 5,
  sim_dist = sim_varf_normal(2, 3, 0, 1, 10)
) %>% unnest(data)

@@

p_varf <- sim_panel_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle(paste("Design $D_{var_f}$:", varf))
```

```{r null}
sim_panel_null <- sim_panel(
   nx = 2,
   nfacet = 3,
   ntimes = 500,
   sim_dist = distributional
   ::dist_normal(0, 1)
) %>% unnest(c(data))

@@

p_null <- sim_panel_null %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle(paste("Design $D_{null}$:", null))
```

```{r plot-all}
library(patchwork)
(p_null + p_varf)/(p_varx + p_varall)
```

The p-value of the design $D_{null}$ is size.
The p-value of other designs is power.
Confidence interval of the

```{r, eval = FALSE, echo = FALSE}
smart_harmony <-read_rds("data/smart_harmony_nonst.rds")
smart_harmony
```

```{r smart_harmony,eval = FALSE, echo = FALSE}

sm <- smart_meter10 %>% dplyr::filter(customer_id %in% c("10017936"))

harmonies <- sm %>% 
  harmony(ugran = "month",
          filter_in = "wknd_wday",
          filter_out = c("hhour", "fortnight"))

harmony_tbl =  harmonies

smart_harmony <- sm %>% 
  rank_harmony(harmony_tbl = harmonies,
               response = "general_supply_kwh", 
               dist_ordered = TRUE)

smart_harmony %>% 
  mutate(MMPD = round(MMPD, 3), max_pd = round(max_pd, 3)) %>% 
  mutate(rankn = row_number()) %>%
  rename("rankun" = "r") %>% kable()
```

The p-value is almost always greater than 0.05, which means there is no significant differences between the different categories, which is true from the simulation design.

```{r size, eval = FALSE}
library(readr)
library(tidyverse)
data_N01 <- read_rds(here("simulations/norm/null_design_quantrans_nperm/data-agg/all_data_wpd_N01.rds"))

data_N01_obs <- data_N01 %>% filter(perm_id==1)
data_N01_samp <- data_N01 %>% filter(perm_id!=1)

#plot_mmpd_null_grid(data_N01_samp, data_N01_obs)

mmpd_dist_null_grid  = data_N01_samp

mmpd_null_orig = data_N01_obs
compute_p_value <- left_join(mmpd_dist_null_grid, mmpd_null_orig,
                               by = c("nx", "nfacet")) %>% 
    group_by(nx, nfacet) %>% 
    summarise(p_value = sum(if_else(abs(value.x)>abs(value.y),1,0))/n(), .groups = 'drop')

  ggplot() + 
    geom_histogram(data = mmpd_dist_null_grid, aes(x = value))  + 
    geom_vline(data = mmpd_null_orig, aes(xintercept = value),colour = "red") +
    geom_text(data = round(compute_p_value, 3), size = 3,  
              aes(x = -Inf,
                  y =  Inf,
                  label = paste("p-value:",p_value),
                  hjust   = 0,
                  vjust   = 1)) + 
    facet_grid(nx ~ nfacet)
```

```{r power, eval = FALSE}
library(readr)
library(tidyverse)
data_N01 <- read_rds(here("simulations/norm/null_design_quantrans_nperm/data-agg/all_data_wpd_N01.rds"))

data_N01_obs <- data_N01 %>% filter(perm_id==1)
data_N01_samp <- data_N01 %>% filter(perm_id!=1)

#plot_mmpd_null_grid(data_N01_samp, data_N01_obs)

mmpd_dist_null_grid  = data_N01_samp

mmpd_null_orig = data_N01_obs
compute_p_value <- left_join(mmpd_dist_null_grid, mmpd_null_orig,
                               by = c("nx", "nfacet")) %>% 
    group_by(nx, nfacet) %>% 
    summarise(p_value = sum(if_else(abs(value.x)>abs(value.y),1,0))/n(), .groups = 'drop')

  ggplot() + 
    geom_histogram(data = mmpd_dist_null_grid, aes(x = value))  + 
    geom_vline(data = mmpd_null_orig, aes(xintercept = value),colour = "red") +
    geom_text(data = round(compute_p_value, 3), size = 3,  
              aes(x = -Inf,
                  y =  Inf,
                  label = paste("p-value:",p_value),
                  hjust   = 0,
                  vjust   = 1)) + 
    facet_grid(nx ~ nfacet)
```

```{r sim_all, echo = FALSE, eval = FALSE}
bind_files = readr::read_rds("../hakear-drake/data/compute_p_value.rds")

bind_files$design = 
  factor(bind_files$design,
         levels = c("null","vary_f", "vary_x", "vary_all"), labels = c("D1", "D2", "D3", "D4"))

ggplot(bind_files %>% 
         filter(distribution == "normal")) +
  geom_point(aes(x=design, y = p_value),
                 alpha = 0.5)+
  facet_grid(nx~nfacet) +
  geom_line(aes(x = design, y = p_value, group = distribution, color = distribution))

ggplot(bind_files %>% 
         filter(distribution == "exponential")) +
  geom_point(aes(x=design, y = p_value),
                 alpha = 0.5)+
  facet_grid(nx~nfacet) +
  geom_line(aes(x = design, y = p_value, group = distribution, color = distribution))

ggplot(bind_files %>% 
         filter(distribution == "gamma")) +
  geom_point(aes(x=design, y = p_value),
                 alpha = 0.5)+
  facet_grid(nx~nfacet) +
  geom_line(aes(x = design, y = p_value, group = distribution, color = distribution))

ggplot(bind_files %>% 
         filter(distribution == "weibull")) +
  geom_point(aes(x=design, y = p_value),
                 alpha = 0.5)+
  facet_grid(nx~nfacet) +
  geom_line(aes(x = design, y = p_value, group = distribution, color = distribution))
```

```{r sim_conf, echo = FALSE, eval = FALSE}
library(dplyr)
library(ggplot2)
ggplot2::ggplot(design_all %>% 
                      filter(design == "D1", distribution == "norm")) +
  geom_boxplot(aes(x = as.factor(nx), 
                   y = mmpd.x), width = 0.1) + 
  facet_wrap(~nfacet) +
  scale_fill_viridis_d() +
  xlab("number of x levels") +
  coord_flip()

@@

Suppose $(A, B) \in H_{N_C}$ be a harmony pair where  $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.
Suppose $q^{i, p}_{A, j}$ denote the quantiles with probability $p$ for the of the $i^{th}$ household for $j^{th}$ category of the cyclic granularity $A$. Similarly, $q^{i, p}_{B, k}$ denotes the same for the $k^{th}$ category of the cyclic granularity $B$. Sample quantiles were computed at $p = 0.01, 0.02, \dots, 0.99$. Jensen-Shannon distances are computed between $q^{i, p}_{A, j}$ and $q^{i, p}_{B, k}$ for each ${j \in J, k \in K}$ to obtain within-facet and between-facet distances. A tuning parameter of $\lambda = 0.67$ has been considered to upweigh the within-facet distances and down-weigh the between facet distances and the maximum of them are obtained to compute $wpd_{raw}$. It is further normalised using the approach described in Section \ref{sec:normalization}. This entire process is repeated for all harmony pairs $(A, B) \in H_{N_C}$ and for each households $i \in i = \{1, 2, 3, 4\}$. The harmony pairs are then arranged in descending order and the important ones with significance level 1%, 5% and 10% are highlighted with `***`, `**` and `*` respectively.

@@

elec_linear <- elec %>% 
  ggplot() +
  geom_line(aes(x = reading_datetime, y = kwh)) +
  facet_wrap(~household_id, nrow = 8, labeller = "label_both",
             strip.position =  "right",
             scales = "free_y") 
#+
 # geom_rect(aes(xmin = as.Date("2018-02-01"), xmax = as.Date("2018-02-28"), ymin = -Inf, ymax = Inf),colour="white", size=0.5, alpha=0.02)

@@

@@

```{r tab-rank-8}
elec_rank <- elec_sig_split %>% 
  select(-c(6,7, 9, 10)) %>% 
  pivot_wider(
              names_from = id,
                values_from = rank) %>% 
  rename("facet variable" = "facet_variable",
         "x variable" = "x_variable",
         "facet levels" = "x_levels",
         "x levels" = "x_levels") %>% 
  slice(1:4)

>>>>

```{r null4by2,fig.cap=" A graphical display with two categories mapped to x-axis and 4 categories mapped to facets with the distribution of a continuous random variable plotted on the y-axis. Display a is not interesting as the distribution of the continuous rv does not depend across x-axis or facet categories. Display b and c are more interesting than a since there is a change in distribution either across facets(b) or x-axis(a). Display d is most interesting as distribution of the rv changes across both facet and x-axis variable."}

sim_varall_normal <- function(nx, nfacet, mean, sd, w) {
  dist_normal((mean + seq(0,
    (nx *
      nfacet - 1),
    by = 1
  ) * w), sd)
}
sim_panel_varall <- sim_panel(
  nx = 2, nfacet = 3,
  ntimes = 500,
  sim_dist = sim_varall_normal(2, 3, 5, 10, 5)
) %>% unnest(data)

p_varall <- sim_panel_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
  ggtitle("")

sim_varx_normal <- function(nx, nfacet, mean, sd, w) {
   rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w), sd), nfacet)
 }

sim_panel_varx <- sim_panel(
   nx = 2, nfacet = 3,
   ntimes = 500,
   sim_dist = sim_varx_normal(2, 3, 5, 10, 5)
 ) %>% unnest(data)

p_varx <- sim_panel_varx %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle("")

sim_varf_normal <- function(nx, nfacet, mean, sd, w) {
  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w), sd), each = nx)
}

sim_panel_varf <- sim_panel(
  nx = 2, nfacet = 3,
  ntimes = 500,
  sim_dist = sim_varf_normal(2, 3, 5, 10, 5)
) %>% unnest(data)

p_varf <- sim_panel_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle("")

sim_panel_null <- sim_panel(
   nx = 2,
   nfacet = 3,
   ntimes = 500,
   sim_dist = distributional
   ::dist_normal(5, 10)
) %>% unnest(c(data))

p_null <- sim_panel_null %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle("")

ggpubr::ggarrange(p_null, p_varf,  p_varx, p_varall, nrow = 2, ncol = 2,
                  common.legend = TRUE,
                  labels = c("a", "b", "c", "d"))

@@

@@

## Algorithm

@@

## Results

```{r varall-new}

@@

# plot
p_varall <- sim_panel_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
  ggtitle(paste("d", round(varall, 2)))
```

```{r varx-new}
sim_varx_normal <- function(nx, nfacet, mean, sd, w) {
   rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w), sd), nfacet)
 }

sim_panel_varx <- sim_panel(
   nx = 2, nfacet = 3,
   ntimes = 500,
   sim_dist = sim_varx_normal(2, 3, 0, 1, 10)
 ) %>% unnest(data)

@@

# plot
p_varx <- sim_panel_varx %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle(paste("c", round(varx, 2)))
```

```{r varf-new}
sim_varf_normal <- function(nx, nfacet, mean, sd, w) {
  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w), sd), each = nx)
}
sim_panel_varf <- sim_panel(
  nx = 2, nfacet = 3,
  ntimes = 500,
  sim_dist = sim_varf_normal(2, 3, 0, 1, 10)
) %>% unnest(data)

@@

p_varf <- sim_panel_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle(paste("b", round(varf, 2)))
```

```{r null-new}
sim_panel_null <- sim_panel(
   nx = 2,
   nfacet = 3,
   ntimes = 500,
   sim_dist = distributional
   ::dist_normal(0, 1)
) %>% unnest(c(data))

@@

p_null <- sim_panel_null %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
    ggtitle(paste("a", round(null, 2)))
```

```{r plot-all-new2}
library(patchwork)
(p_null + p_varf)/(p_varx + p_varall)
```

@@

Suppose $(A, B) \in H_{N_C}$ be a harmony pair where  $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.
Suppose $q^{i, p}_{A, j}$ denote the quantiles with probability $p$ for the of the $i^{th}$ household for $j^{th}$ category of the cyclic granularity $A$. Similarly, $q^{i, p}_{B, k}$ denotes the same for the $k^{th}$ category of the cyclic granularity $B$. Sample quantiles were computed at $p = 0.01, 0.02, \dots, 0.99$. Jensen-Shannon distances are computed between $q^{i, p}_{A, j}$ and $q^{i, p}_{B, k}$ for each ${j \in J, k \in K}$ to obtain within-facet and between-facet distances. A tuning parameter of $\lambda = 0.67$ has been considered to upweigh the within-facet distances and down-weigh the between facet distances and the maximum of them are obtained to compute $wpd_{raw}$. It is further normalized using the approach described in Section \ref{sec:normalization}. This entire process is repeated for all harmony pairs $(A, B) \in H_{N_C}$ and for each households $i \in i = \{1, 2, 3, 4\}$. The harmony pairs are then arranged in descending order and the important ones with significance level 1%, 5% and 10% are highlighted with `***`, `**` and `*` respectively.

@@

elec_linear <- elec %>% 
  ggplot() +
  geom_line(aes(x = date_time, y = kwh)) +
  facet_wrap(~id, nrow = 8, labeller = "label_both",
             strip.position =  "right",
             scales = "free_y") 
elec_linear
#+
 # geom_rect(aes(xmin = as.Date("2018-02-01"), xmax = as.Date("2018-02-28"), ymin = -Inf, ymax = Inf),colour="white", size=0.5, alpha=0.02)

@@

dotplot <- elec_sig_split %>% 
  ggplot() +
  geom_point(aes(x = x_variable,
                 y = facet_variable, 
                 size = wpd, 
                 fill = significant),
             shape=21,
             alpha= 0.8) +
  scale_size_continuous() +
  theme(legend.position = "bottom") +
  scale_fill_viridis(discrete = TRUE, direction = 1) +
  facet_wrap(~id, nrow = 2) +
  guides(fill = guide_legend(), 
         size = guide_legend()) +
  theme_gray(base_size = 12, base_family = "Times") +
   theme(legend.position = "bottom")

@@

```{r tab-rank-8}
elec_rank <- elec_sig_split %>% 
  select(-c(6,7, 9, 10)) %>% 
  pivot_wider(
              names_from = id,
                values_from = rank) %>% 
  rename("facet variable" = "facet_variable",
         "x variable" = "x_variable",
         "facet levels" = "x_levels",
         "x levels" = "x_levels")
